### 2015-12-07 doc/thesis/intro.tex
More thesis writing

Suppose the observed data are the ending configurations of the board, after the
player has concluded a game against an opponent with a known value of $a$. The
model we have designed is very simple, but it is not obvious how to calculate
the likelihood of a particular ending configuration. Indeed, it seems that the
only way is to enumerate every possible path the game could have taken, and
tabulate the ending configurations of each. Clearly, this is infeasible.
Approximate Bayesian computation was designed for situations like these, where
exact likelihoods are not available, perhaps due to the model involving an
algorithm or generative process, but we still want to estimate the parameters.

>>>>

Suppose the observed data are the ending configurations of the board, after the
player has concluded a game against an opponent with a known value of $a$. The
model we have designed is very simple, but it is not obvious how to calculate
the likelihood of a particular ending configuration. Indeed, it seems that the
only way is to enumerate every possible path the game could have taken, and
tabulate the ending configurations of each. Clearly, this is infeasible.
Approximate Bayesian computation was designed for situations like these, where
exact likelihoods are not available, perhaps due to the model involving an
algorithm or generative process.

### 2015-12-07 doc/thesis/methods.tex
More thesis writing

\begin{algorithm}
  \label{alg:nettree}
  \caption{Simulation of an epidemic and transmission tree over a contact network}
  \begin{algorithmic}
    \State infect one node $v$ at random
    \State let $T$ be a single node labeled $1$
    \State $\tip[v] \gets 1$
    \State $n \gets 1$
    \State $t \gets 0$
    \While{$D \neq \emptyset$ and $|I| + |R| < I_{\max}$ and $t < t_{\max}$}
      \State $s \gets \min(t_{\max} - t, \Exponential(\beta + \gamma))$
      \For{$v \in \tip$}
        \State{extend the branch length of $\tip[v]$ by $s$}
      \EndFor
      \State $t \gets t + s$
      \If{$t < t_{\max}$}
        \If{$\Uniform(0, 1) < \nicefrac{\beta}{\beta + \gamma}$}
          \State choose an edge $e = (u, v)$ from $D$ with probability $\beta_e / \beta$
                 and infect $v$
          \State add nodes with labels $n+1$ and $n+2$ to $T$
          \State connect $n+1$ and $n+2$ to $\tip[v]$ in $T$, with branch lengths $0$
          \State $\tip[v] \gets n + 1$
          \State $\tip[u] \gets n + 2$
          \State $n \gets n + 2$
        \Else
          \State choose a node $v$ from $I$ with probability $\gamma_v / \gamma$
                 and remove $v$
          \State remove $v$ from $\tip$
        \EndIf
        \State update $S$, $I$, $R$, $D$, $\beta$, and $\gamma$
      \EndIf
    \EndWhile
  \end{algorithmic}
\end{algorithm}

@@

The tree kernel developed in~\autocite{poon2013mapping} provides a
comprehensive similarity score between two phylogenetic trees. The kernel
computes the dot-product of two feature vectors \ldots

In addition, we implemented a modified version of the normalized
lineages-through-time statistic developed in~\autocite{janzen2015approximate},
which uses piecewise linear functions instead of step functions for the
lineages-through-time plots.

@@

@@

\subsubsection{Identification of separable parameters in kernel space}

>>>>

\begin{algorithm}
  \label{alg:nettree}
  \caption{Simulation of an epidemic and transmission tree over a contact network}
  \begin{algorithmic}
    \State infect one node $v$ at random
    \State let $T$ be a single node with index $1$
    \State $\tip[v] \gets 1$
    \State $n \gets 1$
    \State $t \gets 0$
    \While{$D \neq \emptyset$ and $|I| + |R| < I_{\max}$ and $t < t_{\max}$}
      \State $s \gets \min(t_{\max} - t, \Exponential(\beta + \gamma))$
      \For{$v \in \tip$}
        \State{extend the branch length of $\tip[v]$ by $s$}
      \EndFor
      \State $t \gets t + s$
      \If{$t < t_{\max}$}
        \If{$\Uniform(0, 1) < \nicefrac{\beta}{\beta + \gamma}$}
          \State choose an edge $e = (u, v)$ from $D$ with probability $\beta_e / \beta$
                 and infect $v$
          \State add nodes with labels $n+1$ and $n+2$ to $T$
          \State connect $n+1$ and $n+2$ to $\tip[v]$ in $T$, with branch lengths $0$
          \State $\tip[v] \gets n + 1$
          \State $\tip[u] \gets n + 2$
          \State $n \gets n + 2$
        \Else
          \State choose a node $v$ from $I$ with probability $\gamma_v / \gamma$
                 and remove $v$
          \State remove $v$ from $\tip$
        \EndIf
        \State update $S$, $I$, $R$, $D$, $\beta$, and $\gamma$
      \EndIf
    \EndWhile
  \end{algorithmic}
\end{algorithm}

@@

The tree kernel developed in~\autocite{poon2013mapping} provides a
comprehensive similarity score between two phylogenetic trees. The kernel
computes the dot-product of two feature vectors, corresponding to the two
trees, in the infinite-dimensional feature space of all possible subset trees
with branch lengths. I implemented the fast algorithm developed
in~\autocite{moschitti2006making}, which first enumerates all pairs of subtrees
with the same number of leaf children, and then computes the kernel by dynamic
programming.

In addition, we implemented a modified version of the normalized
lineages-through-time statistic developed in~\autocite{janzen2015approximate},
which uses piecewise linear functions instead of step functions for the
lineages-through-time plots. This modification is to address a potential
inconsistency for trees of different sizes, illustrated in
Figure~\ref{fig:nltt}.

\begin{figure}
  \caption{Comparison of original formulation of normalized
    lineages-through-time, developed in~\autocite{janzen2015approximate}, with
    our modified version using linear interpolation. Here, the red and blue
    trees both have uniformly spaced branching times. Using step functions
    (left), the nLTT of the two trees is non-zero due to the differing numbers
    of internal branches. Using linear interpolation, the nLTT is zero (right).
  The lines on the right graph have been offset for visibility.}
  \includegraphics[scale=0.75]{nltt}
  \label{fig:nltt}
\end{figure}

@@

I implemented the adaptive sequential Monte-Carlo algorithm
developed in~\autocite{del2012adaptive} for approximate Bayesian computation.
In contrast to most existing sequential Monte-Carlo methods, this algorithm
does not require the user to specify a sequence of decreasing tolerances to
approach the target posterior distribution. Rather, the tolerances are computed
adaptively at each step, starting from infinity at the first iteration. The
algorithm may be stopped when the tolerance reaches a user-defined final value,
or when the rate of acceptance of the MCMC kernel reaches a user-defined
threshold. Following a heuristic applied by the
authors~\autocite{del2012adaptive}, we used the latter stopping criterion,
accepting the SMC approximation to the posterior when the MCMC acceptance rate
dropped below 1.5\%.

\subsubsection{Management of computational experiments with combinatorially
expanded parameters}

This reasearch involved many computational experiments performed with all
possible combinations of several parameters. For example, to identify the best
combination of tree kernel parameters (see \ref{subsubsec:kernel} below), we
considered at least three values each of the parameters $\lambda$ and $\sigma$,
corresponding to at least nine possible combinations of these.

@@

\subsubsection{Identification of separable parameters in kernel space}
\label{subsubsec:kernel}
